%\documentclass[handout,space,nooutcomes]{ximera}
\documentclass{ximera}

\input{../preamble.tex}

\title{Sums and products combined, HW 2}
\author{Bart Snapp \and Brad Findell}
\begin{document}
\begin{abstract}
More homework about Sums and Products Combined. 
\end{abstract}
\maketitle

%\emph{Let's set off the following question in a box.} 

\section*{The Context}
\begin{quote}
Suppose you wrote the numbers $1,2,3,\dots,100$ on a chalkboard. You
may take any two numbers, erase them, and then add the sum of their
sum and their product to the list.
\end{quote}

In the first class, we began investigating this context, and we generated a lot of interesting questions.  
In the second class, we addressed some of these questions and made progress on answering some of them.  

\section*{Homework Questions}

\begin{problem}
First note that the process defines a new operation on numbers.  Let's call the operation $\star$.  Given two numbers $a$ and $b$,  $a\star b$ is then ``the sum of their sum and their product.''    In other words, 
\[
a\star b = \answer{a+b+ab}
\]
\end{problem}

\begin{problem}
Is $\star$ a commutative operation?  (Yes or no.) $\answer[format=string]{yes}$

\begin{explanation}
Yes, because $a\star b = \answer{a+b+ab}  = b\star a$.  

\begin{question}
To be more precise, we might write 
\[
a\star b = (a+b)+ab= (b+a)+ba = b\star a,
\]
which demonstrates that the commutativity of $\star$ follows from the commutativity of 
$\answer[format=string]{addition}$ and $\answer[format=string]{multiplication}$. 
\end{question}
\end{explanation}
\end{problem}


\begin{problem}
Is $\star$ an associative operation?  (Yes or no.)  $\answer[format=string]{yes}$

\begin{explanation}
Yes, because 
\[
(a\star b)\star c =  \left(\answer{a+b+ab}\right)\star c = \answer{(a+b+ab)+c + (a+b+ab)c} 
 = a\star \left(\answer{b+c+bc}\right) = a \star(b\star c). 
 \]
%= a + b + c + ab + ac + bc + abc = a+(b+c+bc) + a(b+c+bc) =
\begin{question}
Now that we have established that $\star$ is both associative 
and commutative, the order that we choose numbers \wordChoice{\choice{does} \choice[correct]{does not}} matter.  

For example, 
\begin{align*}
((1\star 2)\star 3 )\star 4) &= (3\star (2\star 1))\star 4) &  \text{by commutativity} \\
                                        &= ((3\star 2)\star 1))\star 4) & \text{by associativity}  \\
                                        &= (4\star (3\star 2)\star 1))) & \text{by associativity}  \\
                                        &= (4\star (2\star 3)\star 1))) & \text{by associativity}  \\
                                        &= (4\star (3\star 2)\star 1))) & \text{by commutativity}  \\
                                        &= ((4\star 3)\star 2)\star 1))) & \text{by associativity},  \\
\end{align*}
which gives some sense of why we can write $1\star 2 \star 3 \star 4$ without parentheses and without ambiguity.  

Thus, in the original situation, if we continue choosing numbers until only one remains, that last number will be 
\[
1\star 2\star 3\star \dots \star 100. 
\]
\end{question}
\end{explanation}
\end{problem}

\begin{problem}
Name an operation that is associative but \textbf{not commutative}.  
\begin{freeResponse}
\begin{hint}
Possible answers: matrix multiplication, function composition.  
\end{hint}
\end{freeResponse}
\end{problem}

\newpage
Back to the situation, let's try some simpler problems.  

\begin{problem}
Try starting with the numbers $1,2,3,\dots, k$ for several small values of $k$.  What do you notice?  

%Beginning with $k = 2$, the pattern is $5, 23, 119, 729, \dots$.  
\[
\begin{array}{c|c}
\text{Beginning List}& \text{Final Value} \\
\hline
1,  2 & \answer{5}  \\
1,  2,  3 & \answer{23} \\  
1,  2,  3 ,  4 & \answer{5!-1} \\
1,  2,  3 ,  4 ,  5 & \answer{6!-1}
\end{array}
\]
\begin{problem}
Try to discern a pattern and make a conjecture about a general formula:  
\[
\begin{array}{c|c}
\text{Beginning List}& \text{Conjectured final Value} \\
\hline
1,  2,  3 ,  \dots, 10 & \answer{11!-1} \\
1,  2,  3 ,  \dots, 100 & \answer{101!-1} \\
1,  2,  3 ,  \dots, n & \answer{(n+1)!-1} 
\end{array}
\]

\begin{hint}
Compare your computed values to factorials.  The answer boxes can interpret expressions with terms such as $4!$ and the like. 
\end{hint}
\end{problem}
\end{problem}

\begin{problem}
Suppose we change the context.  For example, what if we start with just $1$s?  Try starting with $k$ $1$s, for several small values of $k$.  Make a conjecture to generalize the pattern.    

% Beginning with $k=2$, the pattern is $3, 7, 15, 31, \dots$.  

\[
\begin{array}{c|c}
\text{Beginning List}& \text{Final Value} \\
\hline
1,  1 & \answer{3}  \\
1,  1,  1 & \answer{7} \\  
1,  1,  1 ,  1 & \answer{15} \\
1,  1,  1 ,  1 ,  1 & \answer{31} \\
10\ 1s & \answer{2^{10}-1} \\
n\ 1s & \answer{2^n-1} 
\end{array}
\]
\begin{hint}
Compare your computed values to powers of $2$. 
\end{hint}
\end{problem}

\newpage 

\begin{problem}
Expand the expression $(x+a)(x+b)$.  What is the value of this expression when $x=1$?  What do you notice?

\begin{prompt}
\[
(x+a)(x+b) = x^2+ \left(\answer{a+b}\right)x+\answer{ab}.  
\]
When $x=1$, we get $\answer{1+a+b+ab}$, which is $(a\star b)+\answer{1}$.  
\end{prompt}

\end{problem}

\begin{problem}
Expand the expression $(x+a)(x+b)(x+c)$.  What is the value of this expression when $x=1$?  What do you notice?

\begin{prompt}
\[
(x+a)(x+b)(x+c) = x^3+ \left(\answer{a+b+c}\right)x^2+\left(\answer{ab+bc+ac}\right)x+\answer{abc}.  
\]

When $x=1$, we get $\answer{1+a+b+c+ab+bc+ac+abc}$, which is $(a\star b\star c)+\answer{1}$
\end{prompt}
With some care, these ideas can be formalized to prove your conjectured answers to previous problems.  
\end{problem}

\begin{problem}
Compare the related polynomial $(x-a)(x-b)(x-c)$.  (\emph{Note the sign changes.}) The \textbf{roots} of this polynomial are the \textbf{solutions} to the 
equation 
\[
(x-a)(x-b)(x-c)=0, 
\]
and those solutions are $x=\answer{a}$, $x=\answer{b}$, and $x=\answer{c}$.  

\begin{problem}
In this cubic polynomial, the coefficient of $x^2$ is $-(a+b+c)$, and we observe that its opposite, $(a+b+c)$, is the $\answer[format=string]{sum}$ of the roots.  

Furthermore, the constant coefficient is $-(abc)$, and its opposite, $(abc)$, is the $\answer[format=string]{product}$ of the roots.  
\begin{hint}
Think of words like average, product, quotient, sum, difference. 
\end{hint}

\begin{problem}
Correct!  And the coefficient of $x$ is $\answer{ab+bc+ac}$, which is called \textbf{the sum of the roots taken two at a time}.  Note that your expression includes all possible products of two of the roots.  
\newline

With this new language, the sum of the roots taken one at a time is the same as the sum of the roots.  Furthermore, for a cubic polynomial, the sum of the roots taken three at a time is the same as the product of the roots.  This idea generalizes to polynomials of degree $n$, where the coefficient of $x^{n-k}$ is $(-1)^k$ times the sum of the roots taken $k$ at a time.  \textbf{Cool Stuff!!}

And it gets cooler still:  The coefficients of such a polynomial in $x$ are each \textbf{symmetric polynomials} on the roots $a, b, c, \dots$, which is to say that they remain unchanged on any permutation of the roots.  This idea is foundational to \textbf{Galois Theory}, which you may study in a course on Abstract Algebra.  
\end{problem}
\end{problem}
\end{problem}

% Stuff about symmetric polynomials!!

%\begin{problem}
%Can you construct algebraic expressions that will help you solve the previous problems?  
%\begin{freeResponse}
%%Try $(x+1)(x+2)(x+3)\dots(x+k)$ and $(x+1)^k$, again when $x=1$.  
%\end{freeResponse}
%\vfill
%\end{problem}

%\begin{problem}
%Summarize the results above, reminding us what the problem was and
%what the solution is. Explain your reasoning along the way.
%\begin{freeResponse}
%\end{freeResponse}
%\vfill
%\end{problem}

\end{document}
