\documentclass[handout,nooutcomes,space]{ximera}

\input{../preamble.tex}

\title{Math Major Tracks}
\author{Bart Snapp \and Brad Findell}
\begin{document}
\begin{abstract}
There are lots of options for mathematics majors at The Ohio State University. 
\end{abstract}
\maketitle

In the Mathematics Department at The Ohio State University, we offer a
number of tracks for mathematics majors.

\begin{description}
\item[Applied Math Track] Applied Math has two options, the Physics or Chemistry track. The Applied Math concentration is intended for students seeking a strong foundation in mathematics and its application to industrial and physical sciences. Students will be well prepared for a career or graduate-level study in engineering, computer science, physics, architecture and more.
 
\item[Biological Math Track] ``Mathematics is applicable to many different aspects of biology, and new applications are continually being found.'' The Math Biology track is designed to prepare a student for research-based careers and graduate studies in biological and medical fields such as (but not limited to) genomics, neuroscience, and molecular modeling.
 
\item[Educational Math Track] The Education Track is designed for students seeking secondary education licensure in mathematics. Students in the Math Education track will seek teaching certification in a graduate school setting after their undergraduate degree. By learning the foundational logic within a wide scope of math topics from calculus to abstract algebra, this track helps to generate strong educators with exemplary content knowledge as well as the ability to analyze, decipher, and explain math in diverse ways.
 
\item[Financial Math Track] The Financial Math track is designed for students seeking a career in quantitative finance or a related field. Students planning to pursue graduate studies in financial mathematics are strongly encouraged to take the honors versions of the mathematics courses, whenever possible.
 
\item[Theoretical Math Track] The Theoretical Track is intended for students seeking a strong foundation in classical mathematics, in particular for students intending to pursue graduate study in mathematics. Students planning to pursue a PhD. in mathematics are strongly encouraged to take the honors version of the math courses whenever possible.
\end{description}

%\newpage 
%
%\begin{question}[.5in]
%What is the most advanced mathematics course you have completed?    
%\begin{freeResponse}
%\end{freeResponse}
%\end{question}
%
%\begin{question}[.3in]
%Which track (if any) are you in now?  
%\begin{multipleChoice}
%  \choice[correct]{Applied Math Track}
%  \choice[correct]{Biological Math Track}
%  \choice[correct]{Educational Math Track}
%  \choice[correct]{Financial Math Track}
%  \choice[correct]{Theoretical Math Track}
%  \choice[correct]{Not sure/other.}
%\end{multipleChoice}
%\end{question}
%
%\begin{question}[0.3in]
%  Which track might you be interested in?
%  \begin{multipleChoice}
%    \choice[correct]{Applied Math Track}
%    \choice[correct]{Biological Math Track}
%    \choice[correct]{Educational Math Track}
%    \choice[correct]{Financial Math Track}
%    \choice[correct]{Theoretical Math Track}
%    \choice[correct]{Not sure/other.}
%  \end{multipleChoice}
%\end{question}
%

\end{document}
